Generative LLM oversampling for the class imbalance problem in hate speech detection

### 2023-05-13 paper/paper.tex
rewrite discussion and conclusion sections

%P2: Introduce the need for automated methods of hate speech detection
With these developments and the large amount of content on social media sites, these sites have been increasingly looking to automatic detection methods for hate speech \cite{ullmann}. These methods use Machine Learning (ML) to automatically detect and classify hateful speech, removing some of the work done by moderators, whose primary job is to respond reactively to user reports of hate speech \cite{ullmann}. Chandra et al. \cite{chandra} used a combination of image and text processing and classification algorithms to classify images and text on social media sites Twitter and Gab as antisemitic or not \cite{chandra}. In this paper, we seek a lower cost method that uses only the text component of their dataset to further analyze the presence of hate speech on Twitter and investigate new algorithms for classification.

@@

%P4: Wrap up the research question and context
With this in mind, there are two crucial research questions this paper seeks to answer. First: how can automatic ML methods be improved in the domain of text classification for hate speech detection? Second: can generative LLMs be used to allay the class problem in text data? These are the questions we seek to provide answers to, with the hope that they may inform future research and hate speech detection systems.

@@

\section{Discussion}
%P1: talk about difference between binary and 4-class tasks
The most striking result of these trials is the difference in relative performance among the different tasks. Given that a class imbalance exists in all tasks, it may be expected that the generative LLM oversampling would produce similar results for each of these imbalances, but this is not the case. Furthermore, the difference in performance of the generative LLM-trained models to the models trained on data from other resampling methods grows from the 5-class to the 4-class task, suggesting that there is something particular to the binary classification problem that the generative LLM is not as good at accounting for as other resampling methods.

%P2: suggest reasons for this (authenticity of text samples)
One possible explanation could be the size of the classes. In the case of the 4-class classification task, one of the classes has only 8 observations, meaning, during undersampling, all other classes would also be reduced to 8 samples, making a very small dataset, that is perhaps inadequate for classification, providing one possible explanation for why oversampling outperformed undersampling on the 4-class and 5-class problems. As for why the generative LLM specifically outperformed other oversampling methods, it may simply be that generating text and then converting it into a vectorized form produces more authentic results than generating new samples from those vectorized forms with methods such as AdaSyn and SMOTE. This would explain the augmented LLM's outperformance of other oversampling methods on the 4-class and 5-class tasks, but would leave unexplained why it was outperformed by random oversampling in the binary classification task.

\section{Conclusions and Future Work}
%P3: suggest further lines of inquiry (testing of this on images, tabular, testing more texts)
In order to reach a better understanding of the applications of generative LLM oversampling, more research should be done on the abilities of the method to account for class imbalances in text data. Hopefully, future research that applies the method to datasets with differing sizes, levels of severity of imbalance, and number of classes to classify will help pinpoint the exact situations where generative LLM oversampling can be most helpful, informing future systems that may use the method.

This paper proposes the use of generative LLMs to oversample imbalanced text datasets in order to account for the imbalance. Training models with different resampling methods, we find that results are unequal among tasks. The models trained on generative LLM data showed no improvement over other models in the binary classification of social media posts into an antisemitic/not antisemitic dichotomy, but they showed great improvement on the classification of tweets into types of antisemitism. Further research on imbalanced datasets with different characteristics is required to understand why the difference in relative performance was so stark and what factors most contribute to the utility of this method to accounting for the imbalance problem.

\begin{thebibliography}{31}
\bibitem{abdelrahman}
Abd Elrahman, S. M., Abraham, A: A review of class imbalance problem. Journal of Network and Innovative Computing \textbf{1}, 332--340 (2013)

@@

@@

>>>>

%P2: Introduce the need for automated methods of hate speech detection
With these developments and the large amount of content on social media sites, these sites have been increasingly looking to automatic detection methods for hate speech \cite{ullmann}. These methods use Machine Learning (ML) to automatically detect and classify hateful speech, removing some of the work done by moderators, whose primary job is to respond reactively to user reports of hate speech \cite{ullmann}. Chandra et al. \cite{chandra} used a combination of image and text processing and classification algorithms to classify images and text on social media sites Twitter and Gab as antisemitic or not \cite{chandra}. In this paper, we seek a lower cost method that uses only the text component of their dataset to further analyze the presence of hate speech on Twitter and investigate new algorithms for classification.\footnote{Our dataset contains less samples than that of \cite{chandra} because Twitter removed some of the hateful posts before we collected the data.}

@@

%P4: Wrap up the research question and context
With this in mind, along with the recent advances in generative LLMs, we ask whether such methods can allay the class imbalance problem in text data better than traditional approaches, and to what extent that may improve hate speech detection on social media.

@@

\section{Discussion}
The most striking result is the difference in relative performance of the resampling methods between the three tasks. The binary task sees small improvements in recall with overall larger losses in precision, the 5-class task sees large losses in precision with large gains in recall, particularly by the LLM, and the 4-class task sees large gains in recall and precision for the LLM, with other methods having smaller recall gains and performing worse than no resampling on precision. This can primarily be explained through two factors: the degree of imbalance of the data and the degree of appropriateness of an oversampling technique to the task at hand.

First, the degree of imbalance can affect the relative performance of resampling methods. The binary task saw the least benefit from resampling methods, as its losses in precision generally outweighed its gains in recall. This can be explained by the fact that the binary task had relatively large amounts of data for both classes, making resampling less useful compared to other tasks, hence why the observed improvement in recall is less than that of other tasks. For the 5-class task, on the other hand, the classes are imbalanced to a very high degree, with sample sizes ranging from 8 to 877. In order to make these classes equal in size, drastic measures must be taken, either by removing large portions of the dataset or adding a quantity of synthetic or duplicated samples that outnumber the original size of that class within the dataset. This can, in the case of undersampling, remove critical pieces of information that could help the model \cite{marques} \cite{kraiem}, and, in the case of oversampling, introduce excessive amounts of synthetic data or make the dataset prone to overfitting \cite{marques} \cite{kraiem}. This could explain the large decreases in precision in this task. On the 4-class task, the imbalance is far more even, with sample sizes ranging from 8 to 170, certainly still a large difference, but a far cry from the 5-class task. This could explain why the losses in precision for the 4-class task are not as high as those in the 5-class task.

This factor, however, does not explain all of the difference. Why, for example, do the models trained on LLM-generated data outperform other models by such a wide margin in the 4-class task, not only in recall but also precision? Here, it is necessary to look at the methods themselves and the algorithms they use. The generative LLM, being trained on large amounts of text data, has likely been trained partially on samples similar in topic, purpose, or register to the ones present in the dataset. Its familiarity with this type of text allows it to generate more appropriate synthetic samples than those generated purely geometrically by such algorithms as SMOTE and AdaSyn, while avoiding the overfitting that can plague random oversampling as a result of simply duplicating data \cite{marques} \cite{kraiem}.

These two factors, when considered together, can explain a large portion of the observed differences in performance of the resampling methods across the three tasks. The binary case saw very little improvement from any method because the degree of imbalance was not very high, and there was ample data present from both classes. The 5-class task saw the best performance from the models trained on LLM-generated data because this method is the best-suited to the task, but the very large class imbalance negatively affected precision to a high degree. The 4-class task also saw the best performance from these models for the same reason, but the lower degree of class imbalance made the losses in precision lower, making all methods, but especially the generative LLM due to its high appropriateness to the task, perform better on average precision on this task than the previous one.

\section{Conclusions and Future Work}
In this paper, we propose a method of oversampling for text data involving the use of generative LLMs to generate text samples and account for the class imbalance problem. We apply this method to an imbalanced antisemitism classification dataset on three tasks, one binary classification and two multiclass classification. We find that this method does not provide benefits over other methods on the binary classification task, but provides improvements over other methods in the 5-class and especially 4-class tasks. Our findings suggest that generative LLM oversampling, like other resampling methods, does not provide large benefits when there is plentiful data for both classes, and will not perform as well when there is a very large degree of imbalance, but has merit over other methods in certain tasks. We hypothesize that this is because the generative LLMs, by basing the generation on existing text knowledge rather than geometry like other synthetic oversampling methods, are more appropriate to the task of oversampling text data.

For future work in the area, we believe it is important to verify the results and conclusions we have reached. This could be achieved through testing the methods on a wide variety of tasks, with different types of text data (different registers, topics, etc.), different numbers of classes, and different degrees of imbalance. These studies could help test our analyses of the results and pinpoint the situations where generative LLM resampling is most appropriate.

\begin{thebibliography}{33}
\bibitem{abdelrahman}
Abd Elrahman, S. M., Abraham, A: A review of class imbalance problem. Journal of Network and Innovative Computing \textbf{1}, 332--340 (2013)

@@

\bibitem{kraiem}
Kraiem, M. S., Sánchez-Hernández F., Moreno-García M. N.: Selecting the suitable resampling strategy for imbalanced data classification regarding dataset properties. an approach based on association models. Applied Sciences \textbf{11}(18), 8546 (2021)

@@

\bibitem{marques}
Marqués, A. I., García, V., Salvador Sánchez, J.: On the suitability of resampling techniques for the class imbalance problem for credit scoring. Journal of the Operational Research Society \textbf{64}(7), 1060--1070 (2013)
